Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2015-12-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Reviewed conclusions. Some important changes about collocations or readability

In this paper two versions of a client-server architecture for volunteer and distributed
evolutionary algorithms that uses the browser and generated using {\sf
  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
evaluated. Volunteers are {\em in the cloud}, as stated in the title,
since they are a {\em CPU as a service} for the persons running the
experiment. In fact, in this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative or in conjunction with other local or
cloud-based methods to add computing power in a seamless way through
the pool that NodIO creates.

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than the
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over volunteers to other experiments when the
one they were participating on was finished, shown in the low
correlation between the number of IPs in successive experiments. This
also resulted in a low number of generations allotted by users.

@@

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model depends on the implementation, with contributions following
a General Extreme Values distribution in the case of {\sf NodIO} and a
Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
close enough to each other and, in the second case, reflect the fact
that all time spent in the page is actually devoted to computing,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse might be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they've done it
once, they might do it more times. In the spirit of Open Science, this
gamification might involve computing in real time data such as the one
presented in this paper and showing it in the same page.

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar way, one
obtained up to 5 times more total cycles  than the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to explore in the future.

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for other computationally demanding
tasks, especially ones that cover the middle ground between a problem
finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool as was done in \cite{araujo2008mam},
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.

>>>>

In this paper two versions of a client-server architecture for volunteer and distributed
evolutionary algorithms have been evaluated. The architecture uses the browser and has been generated using {\sf
  NodIO}, based on the {\sf NodEO} evolutionary algorithm library. Volunteers are {\em in the cloud}, as stated in the title,
since they are a {\em CPU as a service} for the persons running the
experiment. In fact, in this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative, or in conjunction with other local or
cloud-based methods to add computing power in a seamless way through
the pool that {\sf NodIO} creates.

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than the
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over other experiments to volunteers when the
experiment they were participating on was finished. This has been shown in the low
correlation between the number of IPs in successive experiments, and also resulted in a low number of generations allotted by users.

@@

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model depends on the implementation, with contributions following
a GEV distribution in the case of {\sf NodIO} and a
Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
close enough to each other and, in the second case, reflect the fact
that all time spent in the page is actually devoted to computing,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse might be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they have done it
once, they might do it more times. In the spirit of Open Science, this
gamification might involve computing in real-time data such as the one
presented in this paper, while showing it in the same page.

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. Even if
the three previous experiments were published in a similar way, one
obtained up to 5 times more total cycles  than the one with the least
number of cycles. It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation

Although we think that, in general, the results presented in this
paper are independent of the problem chosen, it is true that even if
it is a difficult problem for evolutionary algorithms, it takes a few
seconds with the right settings to solve. This makes it difficult also
for this kind of asynchronous framework, but it would be interesting
to check the performance for other computationally demanding
tasks, especially ones that cover the middle ground between a problem
finishing in seconds or minutes (Trap) and weeks (Rastrigin F15). 
Besides, the two versions of the framework include as an
algorithmic variant using random population size. We do not think this
has had a big influence on the results and in fact this is not noticed
for F15, which is the second version tested, but it would be interesting to
measure exactly what this influence has been. In general there are
many issues with the evolutionary algorithm implementation itself,
including using different, or adaptive, policies for inserting and
sending individuals to the pool as was done in \cite{araujo2008mam},
using different policies for population initialization, and also the
incorporation of high-speed local resources to the pool to check what
would be the real influence of the volunteer pool to the final
performance.
